---
title: "Project 1- checkpoint 3"
author: "Addie Weaver, Scarlet Kwan, Sophia Woodward"
format: html
editor: source
embed-resources: true
code-tools: true
execute: 
  echo: true
  warning: false
  message: false
---

```{r}

democracy_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-05/democracy_data.csv')
library(tidyverse)
library(gt)
library(glue)
library(httr)
library(tidyjson)
library(jsonlite)
```

## Data Sources

### #1

```{r}
api_key <- "0FimbRUwssZ5BiS0hkHIfUqPXYvtGuFafwlskcZE" 
# link to this data set https://api-ninjas.com/api/gdp

get_gdp <- function(country, key = api_key)
  {
  
  url <- glue("https://api.api-ninjas.com/v1/gdp?country={country};X-Api-Key={api_key}")
  my_data <- safely(fromJSON)(url)
  
  res <- my_data$result
  
  #had to change it to length & null cause i kept getting one or the other
   if (is.null(res) || length(res) == 0) {
    return(tibble(
      country_code = character(),
      year = integer(),
      gdp = double()
    ))
  } else {
    return( #made a tibble for the 3 pieces of data I got from the api
      as_tibble(res) |>
        select(country, year, gdp_nominal) |>
        mutate(
          country_code = country,
          year = as.integer(year),
          gdp = as.numeric(gdp_nominal)
        ) |>
        select(country_code, year, gdp)
    )
  }
}

```

## Data Joins Democracy & GDP For DataSet 1
```{r}


#vector of the country_codes
# get unique country codes only
country_codes <- democracy_data |>
  distinct(country_code) |>
  pull(country_code)

# pull GDP data for all available countries
gdp_df <- country_codes |>
  map(get_gdp) |>
  bind_rows()

#gdp_df

democracy_gdp <- democracy_data |>
  mutate(
    female_leader = is_female_monarch == TRUE | is_female_president == TRUE
  ) |>
  left_join(gdp_df, by = c("country_code", "year"))

democracy_gdp

```

Descript Here

### #2

```{r}
get_unemployment <- function(country, key = api_key)
{
  
  url <- glue("https://api.api-ninjas.com/v1/unemployment?country={country};X-Api-Key={key}")
  my_data <- safely(fromJSON)(url)
  
  res <- my_data$result
  
  if (is.null(res) || length(res) == 0) {
    return(tibble(
      country_code = character(),
      year = integer(),
      unemployment_rate = double()
    ))
  } else {
    return(
      as_tibble(res) |> #made a tibble using the 3 piences of data from api
        select(country, year, unemployment_rate) |>
        mutate(
          country_code = country,
          year = as.integer(year),
          unemployment_rate = as.numeric(unemployment_rate)
        ) |>
        select(country_code, year, unemployment_rate)
    )
  }
}

```
## Data Joins Democracy & Unenployment For DataSet 2
```{r}
country_codes <- democracy_data |>
  distinct(country_code) |>
  pull(country_code)

unemp_df <- country_codes |>
  map(get_unemployment) |>
  bind_rows()

democracy_unemp <- democracy_data |>
  left_join(unemp_df, by = c("country_code", "year"))

democracy_unemp
```


Descript Here


```

## Data Summaries

### \# Research Question: GDP For Top Countries With Most Female Leaders Over Time

```{r}
# Here are top 10 Female Leaders With some things summarized not very nicely but for your reference to do the data summaries or plots
top_10_female_leaders_gdp <- democracy_gdp |>
  filter(!is.na(gdp)) |>
  group_by(country_name, country_code) |>
  summarize(
    avg_gdp = mean(gdp),
    n_female_leaders = sum(female_leader, na.rm = TRUE),
    .groups = "drop"
  ) |>
  slice_max(n_female_leaders, n = 10, with_ties = FALSE)

top_10_female_leaders_gdp

# Some things to consider: We have gdp per year per country, so if you want to do based on years, possible, if you want to do avgs or sums or plots or whatever, all the data is there. This is just the top 10 female leader countries with gdp info.


```

### #2 Research Question: ........

```{r}

```
